---
title: "Visualizing Temperatures from Experimental & Prescribed Fires"
author: "Whalen Dillon"
output: 
  html_document:
    self_contained: no
    toc: yes
---

```{r set options, echo=F}
# knitr::opts_chunk$set(fig.height = 10)
# knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(dpi = 600)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```

```{r load packages}
library(plyr); library(dplyr)
library(readr); library(ggplot2)
```

# Plot Experimental FABIO Burns: Cogongrass & Wiregrass
```{r load data}
# Files are output from 'fire-temps.R' script.
fabio_fires_temperatures <- read_csv("data/fabio-fires-temperatures.csv")
d1 <- read_csv("data/fabio-fires-temperatures.csv")
d2 <- read_csv("data/fabio-fires-biomass.csv")
```

```{r calculate fire temperature metrics}

time_abv100 <- d1 %>% filter(probe_type == "old") %>%
      group_by(location,fire_id) %>%
      filter(tempC>100) %>%
      summarise(s_abv100 = length(tempC)) %>%
      ungroup(.)
      
time_abv300 <- d1 %>% filter(probe_type == "old") %>%
      group_by(location,fire_id) %>%
      filter(tempC>300) %>%
      summarise(s_abv300 = length(tempC)) %>% 
      ungroup(.)

# d1$flame_ht2 <- as.integer(d1$flame_ht2)
max_temp <- d1 %>%
      filter(probe_type == "old") %>%
      group_by(location, fire_id) %>%
      summarise(max_temp = max(tempC)) %>% 
      ungroup(.)

avg_max_temp <- d1 %>% 
      filter(probe_type == "old") %>%
      group_by(location, fire_id) %>% 
      arrange(desc(tempC)) %>% 
      summarise(avg_max_temp = mean(tempC[1:5])) %>% 
      ungroup(.)


d1 <- join_all(list(max_temp, avg_max_temp, time_abv100, time_abv300), 
               type = "full", match = "all")
# summary(d1)
d1$s_abv300[is.na(d1$s_abv300)] <- 0
d1$s_abv100[is.na(d1$s_abv100)] <- 0

d2 <- left_join(d1, d2)
# summary(d2)

# levels(d2$location)
d2$location <- factor(d2$location, levels = c("50cm", "25cm", "0cm"))
d2$fire_id <- as.character(d2$fire_id)

write_csv(d2, "data/experimental-fire-data-summary.csv")

```

## Figures of wiregrass vs. cogongrass fire temperatures
```{r pct fuel moisture comparison}

fm_plot <- ggplot(filter(d2, location == "0cm", pct_green!="NA"), 
       aes(factor(biomass_type), pct_fuel_moisture)) +
      theme_bw() +
      geom_boxplot(aes(fill = factor(pct_green))) +
      scale_fill_discrete(h = c(60,120)) +
      ylab("% fuel moisture") +
      xlab("") +
      labs(fill = "% Green")
fm_plot +
      theme(legend.text = element_text(size = 14),
            axis.text = element_text(size = 16),
            axis.title.y = element_text(size = 16))

      # facet_grid(.~pct_green) +
      # geom_errorbar() +

ggplot(filter(d2, pct_fuel_moisture!="NA", location == "0cm"),
       aes(pct_fuel_moisture, est_pct_fuel_moisture)) +
      geom_point() +
      geom_abline(slope = 1) +
      scale_x_continuous(limits = c(0,61), breaks = seq(0,61,10)) +
      scale_y_continuous(limits = c(0,61), breaks = seq(0,61,10)) +
      theme_bw()
# The estimated fuel moisture higher than the pre-burn fuel moisture, which makes me think that we (I mostly) didn't do a sufficient job of sampling the fuels used in the fire. 

d2 %>% filter(pct_fuel_moisture!="NA", location == "0cm") %>% 
      # select(pct_fuel_moisture, est_pct_fuel_moisture) %>% 
      mutate(diff_pct_fuel_moisture = pct_fuel_moisture - est_pct_fuel_moisture) %>% 
      filter


## Average & Median Fuel Moisture Check
# filter(d2, location == "0cm", pct_green!="NA") %>% 
#       group_by(pct_green, biomass_type) %>% 
#       summarise(avg_fuel_moisture = mean(pct_fuel_moisture,na.rm=T),
#                 med_fuel_moisture = median(pct_fuel_moisture,na.rm=T))

```

Wiregrass has much lower fuel moisture content than cogongrass - with the caveat that these values are based on differences in the weight of the remaining biomass after burning and then drying. In general, we have done a good job of the higher proportion green corresponding to substantially higher fuel moisture within biomass type. Still, there is some overlap of the values for fuel moisture in these groups.

```{r wiregrass-above-100-fuel-moisture}

ggplot(filter(d2, biomass_type=="wiregrass"), 
       aes(factor(standing_biomass), s_abv100)) +
      geom_point(aes(color = pct_fuel_moisture),
                 size = 4) +
      scale_color_gradient(low = "tan", high = "darkgreen") +
      # scale_color_discrete(h = c(60,120)) +
      # geom_line(aes(color = factor(pct_green), linetype = biomass_type)) +
      facet_grid(.~location) +
      ylab("Seconds >100 ºC") +
      xlab("Standing Biomass (g)") +
      theme_bw() +
      theme(legend.position = "right", panel.grid.minor = element_blank()) +
      labs(color = "% Fuel Moisture", 
           linetype = "Fuel\nType",
           shape = "Fuel\nType") +
      ggtitle("Wiregrass Fire Time >100 ºC", subtitle = "No litter")

```

```{r cogongrass-above-100-fuel-moisture}

ggplot(filter(d2, biomass_type=="cogongrass", pct_fuel_moisture!="NA", 
              litter_biomass==0), 
       aes(factor(standing_biomass), s_abv100)) +
      geom_point(aes(color = pct_fuel_moisture),
                 size = 4) +
      scale_color_gradient(low = "tan", high = "darkgreen") +
      # scale_color_discrete(h = c(60,120)) +
      # geom_line(aes(color = factor(pct_green), linetype = biomass_type)) +
      facet_grid(.~location) +
      ylab("Seconds >100 ºC") +
      xlab("Standing Biomass (g)") +
      theme_bw() +
      theme(legend.position = "right", panel.grid.minor = element_blank()) +
      labs(color = "% Fuel Moisture", 
           linetype = "Fuel\nType",
           shape = "Fuel\nType") +
      ggtitle("Cogongrass Fire Time >100 ºC", subtitle = "No litter")

```

```{r pct-consumed-vs-total-biomass}
ggplot(fires_biomass, aes(total_biomass, pct_consumed)) +
      geom_point(aes(color = pct_green, shape = f_litter_biomass),
                 position = "jitter", size = 2) +
      # geom_bar(aes(fill=pct_green), position = "dodge", stat = "identity") +
      theme_bw() +
      facet_grid(litter_biomass ~.)
```

```{r pct-consumed-vs-fuel-moisture}
ggplot(filter(fires_biomass, pct_green!="NA", litter_biomass < 3000),
       aes(pct_fuel_moisture, pct_consumed)) +
      geom_point(aes(color = standing_biomass, shape = pct_green),
                 size = 4) +
      scale_color_gradientn(colours = RColorBrewer::brewer.pal(6, name = "Greens")) +
      theme_bw() +
      facet_grid(litter_biomass ~.)
```


```{r standing biomass no litter cogon vs wire}
ggplot(filter(d2, pct_green>0, litter_biomass==0), 
       aes(standing_biomass, s_abv100)) +
      geom_point(aes(color = factor(pct_green),
                     shape = biomass_type),
                 size = 2) +
      scale_color_discrete(h = c(60,120)) +
      geom_smooth(aes(color = factor(pct_green), linetype = biomass_type), 
                  method = "lm", se = F) +
      # geom_line(aes(color = factor(pct_green), linetype = biomass_type)) +
      facet_grid(.~location) +
      ylab("Seconds >100 ºC") +
      xlab("Standing Biomass (g)") +
      theme_bw() +
      theme(legend.position = "bottom", panel.grid.minor = element_blank()) +
      labs(color = "% Green", 
           linetype = "Fuel\nType",
           shape = "Fuel\nType") +
      ggtitle("Wiregrass & Cogongrass Fire Time >100 ºC",subtitle = "No litter")
# ggsave("figures/standing_wiregrass_cogon_timeabv100.png", dpi = 600)

lm1 <- lm(s_abv100 ~ total_biomass + factor(pct_green) + factor(biomass_type), 
          data = filter(d2, pct_green>0, location=="25cm"))
lm2 <- lm(s_abv100 ~ total_biomass, 
          data = filter(d2, pct_green==75, location=="25cm", biomass_type=="wiregrass"))
lm3 <- lm(s_abv100 ~ total_biomass, 
          data = filter(d2, pct_green==25, location=="25cm", biomass_type=="wiregrass"))
```


```{r total biomass cogon vs wire, fig.width=12, fig.height=9, dpi=600}
ggplot(filter(d2, pct_green>0, location=="0cm"), 
       aes(total_biomass, s_abv100)) +
      geom_point(aes(color = factor(pct_green),
                     shape = biomass_type),
                 size = 4) +
      scale_color_discrete(h = c(60,120)) +
      # geom_line(aes(color = factor(pct_green), 
      #               linetype = biomass_type), size = 1.2) +
      geom_smooth(aes(color = factor(pct_green), linetype = biomass_type), 
                  method = "lm", se = F, size = 1.2) +
      geom_vline(xintercept = 550, linetype = "dotted", size = 1) +
      geom_vline(xintercept = 2070, linetype = "dashed", size = 1) +
      # facet_grid(location~.) +
      ylab("Seconds >100 ºC at cm") +
      xlab(expression(paste("Total Biomass (g/", m^{2}, ")"))) +
      theme_bw() +
      theme(legend.position = c(.8,.2), 
            panel.grid.minor = element_blank(),
            axis.title = element_text(size = 30),
            axis.text = element_text(size = 24),
            title = element_text(size = 30),
            legend.text = element_text(size = 20)) +
      labs(color = "Standing %Green", 
           linetype = "Fuel Type",
           shape = "Fuel Type") +
      ggtitle("Fire Intensity of Standing + Litter Biomass")
# ggsave("figures/total_wiregrass_cogon_timeabv100.png", dpi = 600, width = 12, height = 9)
```

```{r fire profiles wiregrass vs cogongrass, eval=F}
fire_profile1 <- filter(fabio_fires_temperatures, fire_id==1)
# fp1 <- fire_profile1 %>% 
#       group_by(standing_biomass, biomass_type, location, pct_green) %>% 
#       filter(tempC>80) %>%
#       mutate(seconds = seq(1,length(time),1)) %>% 
#       select(tempC,location,standing_biomass,pct_green,biomass_type,seconds)
# fp1$pct_green <- factor(fp1$pct_green, labels = c("25%","75%"))

ggplot(fire_profile1, aes(time, tempC)) +
      geom_line(aes(color = location), size = 1.2) +
      # geom_hline(yintercept = 100, linetype = "F1") +
      # geom_point(aes(shape = biomass_type, color = factor(pct_green)))
      # facet_grid(location~.) +
      theme_bw() +
      xlab("Seconds") +
      ylab("Temperature ºC")# +
      # theme(legend.position = c(.8,.83),
            # axis.title = element_text(size = 30, family = "serif"),
            # axis.text = element_text(size = 24),
            # axis.text = element_text(size = 24, family = "serif"),
            # legend.text = element_text(size = 20, family = "serif"),
            # title = element_text(size = 30, family = "serif")) +
      # labs(color = "%\ Green", linetype = "Fuel Type") +
      # ggtitle("Fire Profiles for Temperatures >80 ºC")
# ggsave("figures/cogongrass-wiregrass-fire-profiles.png", dpi = 600, width = 12, height = 9)
```

```{r wiregrass vs cogongrass max temp, fig.width=10, fig.height=7}
ggplot(filter(d2, pct_green>0, litter_biomass<501), 
       aes(standing_biomass, max_temp)) +
      geom_smooth(method = "lm", aes(color = factor(pct_green), linetype = biomass_type), se = F) +
      geom_point(aes(shape = biomass_type, color = factor(pct_green)), size = 2) +
      # scale_shape_manual(values = c(21, 24)) +
      # scale_linetype_manual(values = c(1,4)) +
      # geom_line(aes(color = factor(pct_green),
      #               linetype = biomass_type)) +
      scale_color_discrete(h = c(60,120)) +
      facet_grid(f_litter_biomass~location) +
      scale_x_continuous(breaks = seq(0,3000,500)) +
      scale_y_continuous(breaks = seq(0,1000,200)) +
      ylab("Maximum Temperature ºC") +
      xlab("Standing Biomass (g)") +
      theme_bw() +
      theme(legend.position = "bottom",
            panel.grid.minor = element_blank(),
            axis.title = element_text(size = 18),
            axis.text = element_text(size = 14),
            title = element_text(size = 20),
            legend.text = element_text(size = 14)) +
      labs(color = "% Green", 
           linetype = "Fuel\nType",
           shape = "Fuel\nType") +
      ggtitle("Wiregrass & Cogongrass Fire Maximum Temperature")
# ggsave("figures/wiregrass_cogon_maxtemp.png", dpi = 300)
```


```{r wiregrass vs cogongrass time above 100, fig.width=10, fig.height=7}
ggplot(filter(d2, pct_green>0, litter_biomass<501), 
       aes(standing_biomass, s_abv100)) +
      geom_point(aes(color = factor(pct_green),
                     shape = biomass_type),
                 size = 2) +
      # geom_line(aes(color = factor(pct_green),
      #               linetype = biomass_type)) +
      geom_smooth(method = "lm", aes(color = factor(pct_green), linetype = biomass_type), se = F) +
      scale_color_discrete(h = c(60,120)) +
      facet_grid(f_litter_biomass~location) +
      scale_x_continuous(breaks = seq(0,3000,500)) +
      scale_y_continuous(breaks = seq(0,1000,100)) +
      ylab("Seconds Above 100 ºC") +
      xlab("Standing Biomass (g)") +
      theme_bw() +
      labs(color = "% Green", 
           linetype = "Fuel\nType",
           shape = "Fuel\nType") +
      theme(legend.position = "bottom",
            panel.grid.minor = element_blank(),
            axis.title = element_text(size = 18),
            axis.text = element_text(size = 14),
            title = element_text(size = 20),
            legend.text = element_text(size = 14)) +
      ggtitle("Wiregrass & Cogongrass Fire Time Above 100 ºC")
```


```{r wiregrass vs cogongrass max temp vs time above 100 ºC}
ggplot(d2 %>% filter(litter_biomass==0),
       aes(s_abv100, max_temp)) +
      geom_point(aes(color = factor(pct_green), 
                     size = standing_biomass, 
                     shape = biomass_type
                     )
                 ) +
      scale_color_discrete(h = c(60,120)) +
      # geom_bar(stat = "identity", position = "dodge", aes(fill = factor(pct_green))) +
      facet_grid(location~.) +
      ylab("Maximum Temperature ºC") +
      xlab("Seconds >100 ºC") +
      theme_bw() +
      scale_size_continuous(breaks = c(200, 1000, 2000, 3000)) +
      theme(legend.position = "bottom", panel.grid.minor = element_blank()) +
      labs(color = "% Green", 
           size = "Standing\nBiomass", 
           shape = "Fuel\nType") +
      ggtitle("Wiregrass & Cogongrass Fire Maximum Temperature vs Time Above 100º C", subtitle = "No litter")
# ggsave("figures/wiregrass_cogon_max_timeabv100.png", dpi = 300)

ggplot(d2 %>% filter(pct_green>0),
       aes(s_abv100, max_temp)) +
      geom_point(aes(color = factor(pct_green), 
                     size = standing_biomass, 
                     shape = biomass_type,
                     alpha = litter_biomass
                     )
                 ) +
      # scale_fill_continuous(low = "yellow", high = "red") +
      # geom_bar(stat = "identity", position = "dodge", aes(fill = factor(pct_green))) +
      facet_grid(location~.) +
      ylab("Maximum Temperature ºC") +
      xlab("Seconds >100 ºC") +
      theme_bw() +
      scale_size_continuous(breaks = c(200, 1000, 2000, 3000)) +
      theme(legend.position = "right", panel.grid.minor = element_blank()) +
      labs(color = "% Green", 
           size = "Standing\nBiomass", 
           shape = "Fuel\nType",
           alpha = "Litter\nBiomass") +
      ggtitle("Wiregrass & Cogongrass Fire Maximum Temperature vs Time Above 100º C")
# ggsave("figures/wiregrass_cogon_max_timeabv100_shading.png", dpi = 300)
```


## Cogongrass Only Figures with Litter Biomass
```{r cogongrass time above 100 ºC}
reg_eqn <- function(df) {
  y = (df$s_abv100)
  x = (df$standing_biomass)
  m <- lm(y ~ x, data = df)
  
  a <- coef(m)[1]
  b <- coef(m)[2]
  r2 <- summary(m)$r.squared
  
  expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
  list(a = format(a, digits = 2),
  b = format(b, digits = 2),
  r2 = format(r2, digits = 3)))
  c(lab = as.character(as.expression(expr)))
}
eqns <- ddply(filter(d2, biomass_type=="cogongrass", litter_biomass < 501, pct_green>0), c("litter_biomass","pct_green","location"), reg_eqn)


ggplot(filter(d2, pct_green>0, biomass_type=="cogongrass", litter_biomass < 501), 
       aes(standing_biomass, s_abv100)) +
      geom_point(aes(shape = factor(litter_biomass),
                     # size = factor(litter_biomass),
                     color = factor(pct_green)
                     # alpha = litter_biomass
                     ), size = 3, position = position_jitterdodge()
                 ) +
      geom_smooth(aes(color = factor(pct_green), linetype = factor(litter_biomass)), method = "lm", se = F, size = 1.2) +
      scale_color_discrete(h = c(60,120)) +
      # geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,  hjust = 0, vjust = -1.5) +
      facet_grid(.~location) +
      scale_x_continuous(breaks = seq(0,max(d2$total_biomass),500)) +
      ylab("Seconds >100 ºC") +
      xlab("Standing Biomass (g)") +
      theme_bw() +
      ggtitle("Cogongrass only") +
      labs(color = "% Green",
           shape = "Litter Biomass (g)",
           linetype = "Litter Biomass (g)",
           size = "Litter\nBiomass (g)"
           ) +
      guides(linetype = guide_legend(override.aes = list(size = .7)),
             color = guide_legend(override.aes = list(linetype = 0))) +
      theme(legend.position = "bottom",
            panel.grid.minor = element_blank(),
            axis.title = element_text(size = 20),
            axis.text = element_text(size = 16),
            title = element_text(size = 20),
            legend.text = element_text(size = 16)) +
      ggtitle("Cogongrass Fire Intensity")
# ggsave("figures/cogon_timeabv100_litter_pctgreen.png", dpi = 300, width = 10, height = 8)

# lm(s_abv100 ~ standing_biomass + litter biomass)
```

```{r cogongrass maximum temperature}
ggplot(filter(d2, pct_green>0, biomass_type=="cogongrass"), 
       aes(standing_biomass, max_temp)) +
      geom_point(aes(color = factor(pct_green),
                     shape = factor(litter_biomass)
                     # alpha = litter_biomass
                     ), size = 2, position = "jitter"
                 ) +
      
      facet_grid(location~.) +
      scale_x_continuous(breaks = seq(0,max(d2$standing_biomass),500)) +
      scale_y_continuous(breaks = seq(0,1400,200)) +
      ylab("Maximum Temperature ºC") +
      xlab("Standing Biomass (g)") +
      theme_bw() +
      theme(legend.position = "bottom", panel.grid.minor = element_blank()) +
      ggtitle("Cogongrass only") +
      labs(color = "% Green",
           shape = "Litter\nBiomass (g)"
           # alpha = "Litter\nBiomass (g)"
           )
ggsave("figures/cogon_maxtemp_shape.png", dpi = 300)
```

```{r cogongrass max temp vs time above 100 ºC}
ggplot(d2 %>% filter(biomass_type=="cogongrass", pct_green>0),
       aes(s_abv100, max_temp)) +
      
      geom_point(aes(size = standing_biomass, 
                     color = factor(pct_green),
                     shape = factor(litter_biomass)
                     # alpha = litter_biomass
                     )
                 ) +
      facet_grid(location~.) +
      ylab("Maximum Temperature ºC") +
      xlab("Seconds >100 ºC") +
      theme_bw() +
      theme(legend.position = "bottom", panel.grid.minor = element_blank()) +
      ggtitle("Cogongrass") +
      labs(color = "% Green",
           size = "Standing\nBiomass (g)",
           # alpha = "Litter\nBiomass (g)"
           shape = "Litter \nBiomass (g)"
           )
# ggsave("figures/cogon_maxtemp_timeabv100_shape.png", dpi = 300)
```

# Plot Prescribed Fire Temperatures

```{r prescribed fire temps}
pfire_temps <- read_csv("data/prescribed-fire-temps.csv")
# unique(pfire_temps$probe_type)
pfire_data <- read_csv("~/Dropbox (UF)/SERDP-Project/data/prescribed-fire-data.csv")
pfire_data <- pfire_data %>%
      filter(date == "7/6/17") %>% 
      rename(standing_biomass=standing_mass_wet,
             litter_biomass=litter_mass_wet
             ) %>% 
      mutate(biomass_type = "native",
             standing_biomass = standing_biomass*16,
             litter_biomass = litter_biomass*16,
             max_flame_ht = apply(cbind(flame_ht1,flame_ht2), 1, function(x) max(x, na.rm = T)),
             max_fuel_ht = apply(cbind(green_ht, brown_ht), 1, max),
             # max_flame_ht = max(c(flame_ht1,flame_ht2),na.rm=T),
             # max_fuel_ht = max(c(green_ht,brown_ht),na.rm=T),
             avg_fuel_ht = rowMeans(cbind(green_ht, brown_ht), na.rm=T),
             avg_flame_ht = rowMeans(cbind(flame_ht1, flame_ht2), na.rm=T)
             ) %>% 
      select(date,id, green_ht:flame_ht2, avg_fuel_ht, avg_flame_ht)
# summary(pfire_data)
# filter(pfire_data, is.na(flame_ht2))
## We had no flame height sensor for the FABIO in the prescribed fire.

pfire_temps_abv100 <- pfire_temps %>% 
      filter(tempC > 100) %>% 
      group_by(location, id) %>% 
      # mutate(heat_flux = tempC*(length(tempC)))
      summarise(s_abv100 = length(tempC),
                heat_flux_abv100 = sum(tempC, na.rm = T))
      
pfire_data_summ <- left_join(pfire_data, pfire_temps_abv100)

pfire_max_temp <- pfire_temps %>%
      group_by(location, id) %>%
      summarise(max_temp = max(tempC, na.rm=T))

pfire_data_summ <- left_join(pfire_data_summ, pfire_max_temp)
pfire_data_summ <- mutate(pfire_data_summ, total_biomass = standing_biomass+litter_biomass)

summary(pfire_data_summ)
filter(pfire_data_summ, is.na(max_temp))
pfire_data_summ <- filter(pfire_data_summ, id!="flame5")
filter(pfire_data_summ, is.na(avg_flame_ht))

# pfire_data_summary <- left_join(pfire_max_temp,pfire_time_abv100)
# pfire_data_summary$s_abv100[is.na(pfire_data_summary$s_abv100)] <- 0
# pfire_data_summ$location <- factor(pfire_data_summ$location, levels = c("50cm", "25cm", "0cm"))
```

```{r pairs plot, eval=F}
source("~/Dropbox (Personal)/R_Stats_resources/panel_cor.R")
pairs(select(filter(pfire_data_summ, location == "50cm"), avg_fuel_ht, avg_flame_ht, time_abv100:total_biomass), lower.panel = panel.cor)
```


```{r flame height vs fuel}
ggplot(filter(pfire_data_summ, id != "fabio naturale"),
       aes(avg_fuel_ht, avg_flame_ht)) +
      geom_point(aes(color = id)) +
      theme_bw()

ggplot(filter(pfire_data_summ, id != "fabio naturale"),
       aes(total_biomass, avg_flame_ht)) +
      geom_point(aes(color = id)) +
      theme_bw()

ggplot(filter(pfire_data_summ, id != "fabio naturale"),
       aes(total_biomass, avg_fuel_ht)) +
      geom_point(aes(color = id)) +
      theme_bw()
```

```{r flame height sensors time above 100 ºC}
# # reg_eqn <- function(df) {
#   y = (df$time_abv100)
#   x = (df$total_biomass)
#   m <- lm(y ~ x, data = df)
#   
#   a <- coef(m)[1]
#   b <- coef(m)[2]
#   r2 <- summary(m)$r.squared
#   
#   expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
#   list(a = format(a, digits = 2),
#   b = format(b, digits = 2),
#   r2 = format(r2, digits = 3)))
#   c(lab = as.character(as.expression(expr)))
# }

# eqns <- ddply(pfire_data_summ, "location", reg_eqn)

ggplot(pfire_data_summ,
       aes(total_biomass, s_abv100, color = id)) +
      geom_point() +
      # geom_smooth(method = "lm") +
      # scale_fill_continuous(low = "yellow", high = "red") +
      # geom_bar(stat = "identity", position = "dodge", aes(fill = factor(pct_green))) +
      # geom_text(aes(x = 450, y = 3, label = lab), data = eqns, parse = T,  hjust = 0, vjust = 0) +
      facet_grid(location~.) +
      # ylab("Seconds >100 ºC") +
      # xlab("Standing Biomass (g/m2)") +
      theme_bw() +
      ggtitle("Prescribed Fire Temperatures from Flame Height Sensors")

ggplot(filter(pfire_data_summ, id != "fabio naturale"),
       aes(avg_flame_ht, s_abv100)) +
      geom_point(aes(shape = location, color = id)) +
      theme_bw()

ggplot(filter(pfire_data_summ, id != "fabio naturale"),
       aes(avg_fuel_ht, s_abv100)) +
      geom_point(aes(shape = location, color = id)) +
      theme_bw()
```

```{r pfire-bar-plots}

pfire_bar_data <- pfire_data_summ %>% 
      group_by(location) %>% 
      summarise(n_obs = length(location),
                avg_s_abv100 = mean(s_abv100),
                se_s_abv100 = sd(s_abv100)/sqrt(n_obs),
                avg_heat_flux = mean(heat_flux_abv100),
                se_heat_flux = sd(heat_flux_abv100)/sqrt(n_obs),
                avg_max_temp = mean(max_temp),
                se_max_temp = sd(max_temp)/sqrt(n_obs))


ggplot(pfire_bar_data,
       aes(location, avg_s_abv100)) +
      geom_bar(stat = "identity", fill = "grey60") +
      # scale_fill_continuous(low = "yellow", high = "red") +
      geom_errorbar(aes(ymin = avg_s_abv100 + se_s_abv100, 
                        ymax = avg_s_abv100 - se_s_abv100),
                    width = .1) +
      # facet_grid(location~.) +
      ylab("Seconds >100 ºC") +
      xlab("Temperature Probe Height") +
      theme_bw() +
      ggtitle(expression(paste("Prescribed Fire Temperatures: biomass 440-670 g/", m^{2})))

ggplot(pfire_bar_data,
       aes(location, avg_max_temp)) +
      geom_bar(stat = "identity", fill = "grey60") +
      # scale_fill_continuous(low = "yellow", high = "red") +
      geom_errorbar(aes(ymin = avg_max_temp + se_max_temp, 
                        ymax = avg_max_temp - se_max_temp),
                    width = .1) +
      # facet_grid(location~.) +
      ylab("Maximum Temperature ºC") +
      xlab("Temperature Probe Height") +
      theme_bw() +
      ggtitle(expression(paste("Prescribed Fire Temperatures: biomass 440-670 g/", m^{2})))

# expression(paste("Basal Area (", m^{2}, ")"))

```

```{r flame height sensors max temp}
ggplot(pfire_data_summ,
       aes(total_biomass, max_temp)) +
      geom_point(aes(color = location)) +
      # scale_fill_continuous(low = "yellow", high = "red") +
      # geom_bar(stat = "identity", position = "dodge", aes(fill = factor(pct_green))) +
      facet_grid(location~.) +
      ylab("Maximum Temperature ºC") +
      # xlab("Standing Biomass (g/m2)") +
      theme_bw() +
      ggtitle("Prescribed Fire Temperatures from Flame Height Sensors")

ggplot(filter(pfire_data_summ, id != "fabio naturale"),
       aes(avg_flame_ht, max_temp)) +
      geom_point(aes(shape = location, color = id)) +
      theme_bw()

ggplot(filter(pfire_data_summ, id != "fabio naturale"),
       aes(avg_fuel_ht, max_temp)) +
      geom_point(aes(shape = location, color = id)) +
      theme_bw()
```

```{r flame height sensor heat flux}
ggplot(pfire_data_summ,
       aes(total_biomass, heat_flux_abv100)) +
      geom_point(aes(shape = location, color = id)) +
      # geom_line(aes(color = location)) +
      # scale_fill_continuous(low = "yellow", high = "red") +
      # geom_bar(stat = "identity", position = "dodge", aes(fill = factor(pct_green))) +
      # facet_grid(location~.) +
      # ylab("Heat Flux >100 ºC") +
      # xlab("Seconds >100 ºC") +
      theme_bw() +
      ggtitle("Prescribed Fire Temperatures from Flame Height Sensors")

ggplot(filter(pfire_data_summ, id != "fabio naturale"),
       aes(avg_flame_ht, heat_flux_abv100)) +
      geom_point(aes(shape = location, color = id)) +
      theme_bw()

ggplot(filter(pfire_data_summ, id != "fabio naturale"),
       aes(avg_fuel_ht, heat_flux_abv100)) +
      geom_point(aes(shape = location, color = id)) +
      theme_bw()
```



# Plot Experimental & Prescribed Fire Temperatures

```{r fabio and prescribed fire, eval=F}
# levels(d2$location)
pfire_data_summ$pct_green[pfire_data_summ$pct_green<50] <- 25
names(d2)
pfire_data_summ <- pfire_data_summ %>% 
      rename(fire_id = id, s_abv100 = s_abv100)


d3 <- rbind(d2,pfire_data_summ)
d3$fire_type <- "experimental"
d3$fire_type[d3$biomass_type=="native"] <- "prescribed"
d3$biomass_type[d3$biomass_type=="wiregrass"] <- "native"
d3$location <- factor(d3$location, levels = c("50cm", "25cm", "0cm"))
d3$total_biomass = d3$standing_biomass+d3$litter_biomass

ggplot(filter(d3, pct_green==25), 
       aes(s_abv100, max_temp, color = litter_biomass)) +
      geom_point(aes(size = standing_biomass, shape = biomass_type)) +
      # scale_fill_continuous(low = "yellow", high = "red") +
      # geom_bar(stat = "identity", position = "dodge", aes(fill = factor(pct_green))) +
      facet_grid(location~pct_green) +
      ylab("Maximum Temperature ºC") +
      xlab("Seconds >100 ºC") +
      theme_bw() +
      ggtitle("Fire Temperatures from Experimental & Prescribed Fires")

ggplot(filter(d3, litter_biomass<=500, pct_green==25, standing_biomass<1000, location == "25cm"),
       aes(standing_biomass, s_abv100)) +
      geom_point(aes(shape = fire_type, color = biomass_type), position = "jitter") +
      # scale_fill_continuous(low = "yellow", high = "red") +
      # geom_bar(stat = "identity", position = "dodge", aes(fill = factor(pct_green))) +
      # facet_grid(factor(pct_green)~.) +
      ylab("Seconds >100 ºC") +
      xlab("Standing Biomass (g)") +
      theme_bw() +
      ggtitle("Fire Temperatures from Experimental & Prescribed Fires")

ggplot(filter(d3, litter_biomass<=500, between(standing_biomass,1,1000)), 
       aes(standing_biomass, max_temp, color = factor(pct_green))) +
      geom_point(aes(shape = biomass_type)) +
      # scale_fill_continuous(low = "yellow", high = "red") +
      # geom_bar(stat = "identity", position = "dodge", aes(fill = factor(pct_green))) +
      facet_grid(location~fire_type) +
      ylab("Maximum Temperature ºC") +
      xlab("Standing Biomass (g)") +
      theme_bw() +
      ggtitle("Fire Temperatures from Experimental & Prescribed Fires")

ggplot(filter(d3, litter_biomass<=500, between(standing_biomass,1,1000)), 
       aes(total_biomass, max_temp, color = factor(pct_green))) +
      geom_point(aes(shape = biomass_type)) +
      # scale_fill_continuous(low = "yellow", high = "red") +
      # geom_bar(stat = "identity", position = "dodge", aes(fill = factor(pct_green))) +
      facet_grid(location~fire_type) +
      ylab("Maximum Temperature ºC") +
      xlab("Total Biomass (g)") +
      theme_bw() +
      ggtitle("Fire Temperatures from Experimental & Prescribed Fires")
```

